---
title: "R Notebook"
output: html_notebook
---

1. There is a relationship between the buying price and car acceptability.
2. Safety ratings significantly impact car acceptability.
3. Maintenance cost influences the overall evaluation of cars.
4. Relationship between the Number of Doors and Car Acceptability.
5. Association between Safety and Maintenance Costs
6. Correlation between Buying Price and Technical Characteristics.
7. Influence of the Number of Persons on Car Acceptability.
8. Analysis of Acceptability Across Different Maintenance and Safety Levels.
9. Identification of Ideal Car Characteristics for 'Good' and 'Very Good' Ratings.
10. Impact of Luggage Boot Size on Car Acceptability
```{r}
columns <-c("buying","maint","doors","persons","lug_boot","safety","class")

df <- read.csv('data/car.data',col.names = columns)
```

```{r}
# Display the first few rows of the dataset
head(df)
```
```{r}
# Check for missing values
summary(is.na(df))
```

```{r}
# Explore basic statistics of numerical attributes
summary(df)

```


```{r}
# Explore unique values in categorical attributes
sapply(df, function(x) length(unique(x)))
```


```{r}
# Explore class distribution
table(df$class)

```

```{r}
# Data Visualization 
library(ggplot2)
```

```{r}
# Scatterplot to visualize relationships between numerical variables
ggplot(df, aes(x = buying, y = maint, color = class)) +
  geom_point() +
  ggtitle("Relationship between Buying and Maintenance Costs") +
  theme_minimal()

```

```{r}

# Barplot to visualize class distribution
ggplot(df, aes(x = class)) +
  geom_bar() +
  ggtitle("Class Distribution") +
  theme_minimal()
```


```{r}

# Convert 'buying' and 'class' to ordered factors
df$buying <- factor(df$buying, levels=c('low', 'med', 'high', 'vhigh'), ordered=TRUE)
df$class <- factor(df$class, levels=c('unacc', 'acc', 'good', 'vgood'), ordered=TRUE)

# Boxplot for buying price vs. car acceptability
boxplot(class ~ buying, data = df, order = c('low', 'med', 'high', 'vhigh'),
        main = 'Relationship between Buying Price and Car Acceptability',
        xlab = 'Buying Price', ylab = 'Car Acceptability')


```


```{r}
# count plotst for safety vs. car acceptability
contingency_table <- table(df$safety, df$class)
barplot(contingency_table, beside=TRUE, col=rainbow(length(unique(df$class))), legend=rownames(contingency_table),
        main='Impact of Safety Ratings on Car Acceptability', xlab='Safety', ylab='Count')

```


```{r}
# Visualization of maintenance vs. car acceptability
ggplot(df, aes(x=maint, fill=class, order=maint)) +
  geom_bar(position='stack') +
  labs(title='Influence of Maintenance Cost on Car Acceptability') +
  theme_minimal()

```

```{r}
# Stacked bar chart for number of doors vs. car acceptability
contingency_table_doors <- table(df$doors, df$class)
barplot(contingency_table_doors, beside=TRUE, col=rainbow(length(unique(df$class))), legend=rownames(contingency_table_doors),
        main='Relationship between Number of Doors and Car Acceptability', xlab='Number of Doors', ylab='Count')

```

```{r}
# Convert categorical variables to numeric
df$doors <- as.numeric(factor(df$doors))
df$persons <- as.numeric(factor(df$persons))
df$lug_boot <- as.numeric(factor(df$lug_boot))

# Scatterplot matrix with different colors for 'buying'
pairs(df[, c("doors", "persons", "lug_boot")], col = as.numeric(factor(df$buying)))

```



```{r}
# Stacked bar chart for number of persons vs. car acceptability
contingency_table_persons <- table(df$persons, df$class)
barplot(contingency_table_persons, beside=TRUE, col=rainbow(length(unique(df$class))), legend=rownames(contingency_table_persons),
        main='Influence of Number of Persons on Car Acceptability', xlab='Number of Persons', ylab='Count')

```






```{r}
# Filter dataset for 'good' and 'v-good' ratings
high_rated_cars <- subset(df, class %in% c('good', 'v-good'))

# Display characteristics of high-rated cars
summary(high_rated_cars)

```






```{r}
# Stacked bar chart for luggage boot size vs. car acceptability
contingency_table_boot <- table(df$lug_boot, df$class)
barplot(contingency_table_boot, beside=TRUE, col=rainbow(length(unique(df$class))), legend=rownames(contingency_table_boot),
        main='Impact of Luggage Boot Size on Car Acceptability', xlab='Luggage Boot Size', ylab='Count')

```












